---
title: "R Notebook"
output: html_notebook
---


```{r}
library(lme4)
library(ggplot2)
library(plotrix)
library(visreg)
library(stringr)
library(tidyverse)
library(ggpubr)
library(gridExtra)
library(tidyverse)
library(tidymodels)
library(vip) 
library(plm)
```


```{r}
new_format_lasso_data<- bind_rows(read_tsv("/Users/noa/Workspace/lasso_positions_sampling_results/f_lasso60.tsv"),
                                  read_tsv("/Users/noa/Workspace/lasso_positions_sampling_results/f_lasso30.tsv"),
                                  read_tsv("/Users/noa/Workspace/lasso_positions_sampling_results/f_lasso15.tsv")
                                  )
new_format_lasso_data<- new_format_lasso_data %>% 
           separate(dataset_id, "_supermatrices_",
                into = c("prefix", "last_value"), 
                remove = FALSE) %>% separate(last_value,"_",into = c("dataset_name", "suffix")) %>%  select(-c("prefix","suffix"))
```

Get a list of all relevant datasets
```{r}
new_format_lasso_data %>% distinct (dataset_name, MSA_original_n_seq, actual_n_seq, job_ind) %>% filter (actual_n_seq==30)
```




Example MSA:
Results are based on:
/groups/pupko/noaeker/lasso_positions_sampling_results/f2_mg30/job_0/_groups_pupko_noaeker_data_supermatrices_edited__supermatrices_NagyA1_NagyA1.fasta/n_seq_30/n_loci_80000/Lasso_folder/exponential/training_4000_random_tree_eval/trimmed_4000/threshold_0.05_outputs/training_sitelh_df_prediction.csv

/groups/pupko/noaeker/lasso_positions_sampling_results/f2_mg30/job_0/_groups_pupko_noaeker_data_supermatrices_edited__supermatrices_NagyA1_NagyA1.fasta/n_seq_30/n_loci_80000/Lasso_folder/exponential/training_4000_random_tree_eval/trimmed_4000/threshold_0.05_outputs/test_sitelh_df_prediction_no_opt.csv

```{r}
example_msa_data<- new_format_lasso_data %>% filter (dataset_name=="NagyA1", actual_n_seq==30, actual_n_loci == 80000)

example_msa_data_specific_configuration<- example_msa_data %>% filter (actual_sample_pct==0.05, actual_training_size==4000)

example_msa_data_specific_configuration %>% select (alpha, number_loci_chosen, `lasso_test_R^2`, lasso_test_spearmanr )

example_msa_training_ll<- read_csv("/Users/noa/Workspace/lasso_positions_sampling_results/NagyA1_training.csv")

g01<-ggplot(example_msa_test_ll,aes(true_test_ll, predicted_test_ll)) +theme_classic()+
  geom_smooth(method='lm')+xlab("True LL")+ylab("Predicted LL ")+geom_point()+ labs(title="")+theme(plot.title = element_text(hjust = 0.5))+ theme(plot.title = element_text(hjust = 0.5))+ theme(text = element_text(size = 15))

g02<-ggplot(example_msa_training_ll,aes(true_training_ll, predicted_training_ll)) +theme_classic()+
  geom_smooth(method='lm')+xlab("True LL")+ylab("Predicted LL ")+geom_point()+ labs(title="")+theme(plot.title = element_text(hjust = 0.5))+ theme(plot.title = element_text(hjust = 0.5))+theme(text = element_text(size = 15))

figure_1<-ggarrange(g02, g01, hjust=-1,vjust=1, heights=c(2,2),common.legend=TRUE,
          labels = c("A", "B"),
          ncol = 1, nrow =2)
figure_1

ggsave("Figure_1.jpeg",
       plot = figure_1, dpi = 350)

dev.off()


r2.taining<- cor.test(example_msa_training_ll$predicted_training_ll,example_msa_training_ll$true_training_ll )
r2.test<- cor.test(example_msa_test_ll$predicted_test_ll,example_msa_test_ll$true_test_ll )

r2.taining
r2.test

r2.taining.spearman<- cor.test(example_msa_training_ll$predicted_training_ll,example_msa_training_ll$true_training_ll, method = "spearman" )
r2.test.spearman<- cor.test(example_msa_test_ll$predicted_test_ll,example_msa_test_ll$true_test_ll, method = "spearman" )

r2.taining.spearman
r2.test.spearman

train_errors<-example_msa_training_ll %>% mutate(error_pct = round(abs((true_training_ll-predicted_training_ll)/true_training_ll)*100,3))
train_errors %>% head(5)
test_errors<-example_msa_test_ll %>% mutate(error_pct = round(abs((true_test_ll-predicted_test_ll)/true_test_ll)*100,3)) 
test_errors %>% head (5)
train_errors %>% summarise(std_error_pct_train = var(error_pct), mean_error_pct_train = mean(error_pct))
test_errors %>% summarise(std_error_pct_test = var(error_pct), mean_error_pct_test = mean(error_pct))


```

Explore unexplained variance on example MSA
```{r}
example_msa_data <- example_msa_data %>% mutate(unexplained_var =1-`lasso_test_R^2`)

example_msa_data %>% select (actual_training_size, actual_sample_pct,`lasso_test_R^2`, unexplained_var)

figure_2<-example_msa_data %>% ggplot(aes(x=actual_sample_pct,y=unexplained_var, group = as.factor(actual_training_size), colour = as.factor(actual_training_size)))  + geom_point(size =3)+ geom_line()+ scale_x_continuous(breaks = c(0.01,0.025,0.05,0.1) ,labels = scales::percent) +labs(color="Training size",x="Sample percentage",y="Unexplained variance (%)")  + scale_y_continuous(labels = scales::percent)+theme_classic() + theme(text = element_text(size = 15)) 


ggsave("Figure_2.jpeg",
       plot = figure_2, dpi = 350)


```

```{r}
supp_data2<- example_msa_data %>% select (actual_training_size, actual_sample_pct,number_loci_chosen, unexplained_var)

supp_data2 %>% write_csv("/Users/noa/Workspace/lasso_positions_sampling_results/supp2.csv")
```



r4s on one msa:
Based on:
1./groups/pupko/noaeker/lasso_positions_sampling_results/f_lasso30/job_0/_groups_pupko_noaeker_data_supermatrices_edited__supermatrices_NagyA1_NagyA1.fasta/n_seq_30/n_loci_80000/r4s.res /Users/noa/Workspace/lasso_positions_sampling_results/r4s_NAGY_res.full

2. /groups/pupko/noaeker/lasso_positions_sampling_results/f2_mg30/job_0/_groups_pupko_noaeker_data_supermatrices_edited__supermatrices_NagyA1_NagyA1.fasta/n_seq_30/n_loci_80000/Lasso_folder/exponential/training_4000_random_tree_eval/trimmed_4000/lasso_rate_4_site /Users/noa/Workspace/lasso_positions_sampling_results/r4s_NAGY_res.lasso

```{r}

full_r4s<- read_csv("/Users/noa/Workspace/lasso_positions_sampling_results/NagyA1_full_rate4site")
lasso_r4s<- read_csv("/Users/noa/Workspace/lasso_positions_sampling_results/NagyA1_lasso_rate4site")
full_r4s<- full_r4s %>% mutate(data = "All MSA positions")
full_r4s
lasso_r4s<- lasso_r4s %>% mutate(data = "Positions chosen by Lasso")
lasso_r4s
all_r4s_data<-rbind(full_r4s,lasso_r4s)
all_r4s_data
figure_3<-all_r4s_data %>% ggplot(aes(x = Numbers, fill = data)) + 
geom_histogram(bins=8,aes(y=..density..),
                 alpha=0.5,position='identity',binwidth=0.5) + labs(fill = "", x = "Standartaized rates", y= "Density") +theme_classic()+ theme(text = element_text(size = 15)) 

summary(lasso_r4s %>% pull(Numbers))

ttest<-t.test(lasso_r4s %>% pull(Numbers),full_r4s%>%pull(Numbers))

ggsave("Figure_3.jpeg",
       plot = figure_3, dpi = 350)



```


Explore unexplained variance on all MSAs

```{r}
supp_data2<- new_format_lasso_data %>% select(dataset_name,actual_n_seq, actual_n_loci,actual_training_size, actual_sample_pct, number_loci_chosen ,`lasso_test_R^2`) %>% arrange(dataset_name)

supp_data2 %>% write_csv("/Users/noa/Workspace/lasso_positions_sampling_results/supp2.csv")


new_format_lasso_data  %>%  mutate (unexplained_var = 1-`lasso_test_R^2`) %>% 
  filter (actual_n_loci==80000) %>%
  ggplot(aes(x=actual_sample_pct,y=unexplained_var, group = interaction(actual_training_size,dataset_name), color = as.factor(actual_training_size)))+ geom_line() + geom_point(size=1.5) + scale_x_continuous(labels = scales::percent,breaks = c(0.01,0.025,0.05,0.1)) +labs(color="Training size" ,x="Sample percentage",y="Unexplained variance on test set (%)") + theme_classic()+facet_grid(cols = vars(dataset_name), rows=  vars(actual_n_seq)) + theme(axis.text.x = element_text(angle = 60, vjust = 0.9, hjust=1))


```

New lme analysis

```{r}
library(nlme)
library(rsq)

lasso_glm_data<- new_format_lasso_data  %>% filter (actual_training_size==4000, actual_sample_pct==0.05) %>% select (`lasso_test_R^2`,actual_n_seq,actual_n_loci,dataset_name,divergence, constant_sites_pct, lasso_mean_rate4site,avg_entropy,alpha ) %>% mutate (unexplained_var =1-`lasso_test_R^2` )
lasso_glm_data
lasso_mixed = lme(log(unexplained_var)/(1-unexplained_var) ~ (actual_n_loci+actual_n_seq+divergence+constant_sites_pct+alpha ), random = ~ 1 | dataset_name, data = lasso_glm_data)
print(summary(lasso_mixed))
print(rsq(lasso_mixed))
plot.lme((lasso_mixed))

```







